---
title: "Real-World Tower of Hanoi - Comprehensive Analysis"
author: "Research Team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 6)

# Load required libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
library(gridExtra)
library(car)  # For ANOVA diagnostics
library(emmeans)  # For estimated marginal means
library(interactions)  # For interaction plots
```

# Executive Summary

This R Markdown document provides a comprehensive statistical analysis of the Real-World Tower of Hanoi experimental study. The experiment investigates how environmental alterations and different resolution strategies affect the solvability and efficiency of the classic Tower of Hanoi problem.

## Research Questions

1. How does disk count (Factor A) affect solvability and moves required?
2. How does alteration rate (Factor B) impact performance?
3. Which illegal formation handler (Factor C) is most effective?
4. Which ground handler (Factor D) minimizes moves?
5. Which duplicate handler (Factor E) performs better?
6. What interactions exist between factors?

---

# Data Loading and Preparation

```{r load_data}
# Load the experimental results
data <- read_csv("data/results.csv", col_types = cols(
  run_id = col_character(),
  run_timestamp = col_character(),
  disk_count = col_integer(),
  target_alteration_percent = col_integer(),
  illegal_formation_handler = col_character(),
  ground_handler = col_character(),
  duplicate_handler = col_character(),
  is_solvable = col_logical(),
  total_moves_to_solve = col_integer(),
  total_alterations = col_integer(),
  total_illegal_states = col_integer(),
  actual_alteration_percent = col_double(),
  raw_move_log_path = col_character()
))

# Display first few rows
head(data) %>%
  kable(caption = "Sample of Raw Experimental Data") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

```{r data_summary}
# Basic data summary
cat("Total number of simulation runs:", nrow(data), "\n")
cat("Successful runs:", sum(data$is_solvable, na.rm = TRUE), "\n")
cat("Failed runs:", sum(!data$is_solvable, na.rm = TRUE), "\n")
cat("Success rate:", round(mean(data$is_solvable) * 100, 2), "%\n")
```

```{r factor_conversion}
# Convert factors to factors (currently characters)
data <- data %>%
  mutate(
    disk_count = factor(disk_count),
    target_alteration_percent = factor(target_alteration_percent),
    illegal_formation_handler = factor(illegal_formation_handler),
    ground_handler = factor(ground_handler),
    duplicate_handler = factor(duplicate_handler)
  )

# Create a filtered dataset with only solvable runs for move analysis
data_solvable <- data %>% filter(is_solvable == TRUE)

cat("Runs available for move analysis (solvable only):", nrow(data_solvable), "\n")
```

---

# Descriptive Statistics

## Factor Levels

```{r factor_levels}
# Display all factor levels
factor_summary <- data.frame(
  Factor = c("A - Disk Count", "B - Alteration %", "C - Formation Handler", 
             "D - Ground Handler", "E - Duplicate Handler"),
  Levels = c(
    paste(levels(data$disk_count), collapse = ", "),
    paste(levels(data$target_alteration_percent), collapse = ", "),
    paste(levels(data$illegal_formation_handler), collapse = ", "),
    paste(levels(data$ground_handler), collapse = ", "),
    paste(levels(data$duplicate_handler), collapse = ", ")
  )
)

factor_summary %>%
  kable(caption = "Experimental Factors and Their Levels") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Success Rate by Factor

```{r success_rate_by_factor}
# Calculate success rate for each factor level
success_by_disk <- data %>%
  group_by(disk_count) %>%
  summarise(
    n = n(),
    successes = sum(is_solvable),
    success_rate = mean(is_solvable) * 100
  )

success_by_alteration <- data %>%
  group_by(target_alteration_percent) %>%
  summarise(
    n = n(),
    successes = sum(is_solvable),
    success_rate = mean(is_solvable) * 100
  )

success_by_formation <- data %>%
  group_by(illegal_formation_handler) %>%
  summarise(
    n = n(),
    successes = sum(is_solvable),
    success_rate = mean(is_solvable) * 100
  )

success_by_ground <- data %>%
  group_by(ground_handler) %>%
  summarise(
    n = n(),
    successes = sum(is_solvable),
    success_rate = mean(is_solvable) * 100
  )

success_by_duplicate <- data %>%
  group_by(duplicate_handler) %>%
  summarise(
    n = n(),
    successes = sum(is_solvable),
    success_rate = mean(is_solvable) * 100
  )

# Display tables
success_by_disk %>%
  kable(caption = "Success Rate by Disk Count", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

success_by_alteration %>%
  kable(caption = "Success Rate by Alteration Percentage", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

success_by_formation %>%
  kable(caption = "Success Rate by Formation Handler", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

success_by_ground %>%
  kable(caption = "Success Rate by Ground Handler", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

success_by_duplicate %>%
  kable(caption = "Success Rate by Duplicate Handler", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Moves Required by Factor (Solvable Runs Only)

```{r moves_by_factor}
# Calculate mean moves for solvable runs
moves_by_disk <- data_solvable %>%
  group_by(disk_count) %>%
  summarise(
    n = n(),
    mean_moves = mean(total_moves_to_solve, na.rm = TRUE),
    sd_moves = sd(total_moves_to_solve, na.rm = TRUE),
    median_moves = median(total_moves_to_solve, na.rm = TRUE)
  )

moves_by_alteration <- data_solvable %>%
  group_by(target_alteration_percent) %>%
  summarise(
    n = n(),
    mean_moves = mean(total_moves_to_solve, na.rm = TRUE),
    sd_moves = sd(total_moves_to_solve, na.rm = TRUE),
    median_moves = median(total_moves_to_solve, na.rm = TRUE)
  )

moves_by_formation <- data_solvable %>%
  group_by(illegal_formation_handler) %>%
  summarise(
    n = n(),
    mean_moves = mean(total_moves_to_solve, na.rm = TRUE),
    sd_moves = sd(total_moves_to_solve, na.rm = TRUE),
    median_moves = median(total_moves_to_solve, na.rm = TRUE)
  )

# Display tables
moves_by_disk %>%
  kable(caption = "Moves Required by Disk Count", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

moves_by_alteration %>%
  kable(caption = "Moves Required by Alteration Percentage", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

moves_by_formation %>%
  kable(caption = "Moves Required by Formation Handler", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

# Visualizations

## Success Rate Visualizations

```{r success_rate_plots, fig.height=8}
p1 <- ggplot(success_by_disk, aes(x = disk_count, y = success_rate, fill = disk_count)) +
  geom_col() +
  geom_text(aes(label = paste0(round(success_rate, 1), "%")), vjust = -0.5) +
  labs(title = "Success Rate by Disk Count", x = "Disk Count", y = "Success Rate (%)") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(success_by_alteration, aes(x = target_alteration_percent, y = success_rate, fill = target_alteration_percent)) +
  geom_col() +
  geom_text(aes(label = paste0(round(success_rate, 1), "%")), vjust = -0.5) +
  labs(title = "Success Rate by Alteration %", x = "Target Alteration %", y = "Success Rate (%)") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(success_by_formation, aes(x = illegal_formation_handler, y = success_rate, fill = illegal_formation_handler)) +
  geom_col() +
  geom_text(aes(label = paste0(round(success_rate, 1), "%")), vjust = -0.5) +
  labs(title = "Success Rate by Formation Handler", x = "Handler", y = "Success Rate (%)") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(success_by_ground, aes(x = ground_handler, y = success_rate, fill = ground_handler)) +
  geom_col() +
  geom_text(aes(label = paste0(round(success_rate, 1), "%")), vjust = -0.5) +
  labs(title = "Success Rate by Ground Handler", x = "Handler", y = "Success Rate (%)") +
  theme_minimal() +
  theme(legend.position = "none")

grid.arrange(p1, p2, p3, p4, ncol = 2)
```

## Moves Distribution

```{r moves_distribution}
ggplot(data_solvable, aes(x = total_moves_to_solve)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Total Moves (Solvable Runs)", 
       x = "Total Moves", y = "Frequency") +
  theme_minimal()
```

## Moves by Disk Count

```{r moves_by_disk_boxplot}
ggplot(data_solvable, aes(x = disk_count, y = total_moves_to_solve, fill = disk_count)) +
  geom_boxplot() +
  labs(title = "Moves Required by Disk Count", 
       x = "Disk Count", y = "Total Moves") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Moves by Alteration Rate

```{r moves_by_alteration_boxplot}
ggplot(data_solvable, aes(x = target_alteration_percent, y = total_moves_to_solve, fill = target_alteration_percent)) +
  geom_boxplot() +
  labs(title = "Moves Required by Alteration Percentage", 
       x = "Target Alteration %", y = "Total Moves") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Interaction Plot: Disk Count × Alteration Rate

```{r interaction_disk_alteration}
interaction_data <- data_solvable %>%
  group_by(disk_count, target_alteration_percent) %>%
  summarise(mean_moves = mean(total_moves_to_solve, na.rm = TRUE), .groups = "drop")

ggplot(interaction_data, aes(x = target_alteration_percent, y = mean_moves, 
                              color = disk_count, group = disk_count)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(title = "Interaction: Disk Count × Alteration Rate",
       x = "Target Alteration %", y = "Mean Total Moves",
       color = "Disk Count") +
  theme_minimal()
```

## Illegal States Encountered

```{r illegal_states}
ggplot(data, aes(x = disk_count, y = total_illegal_states, fill = disk_count)) +
  geom_boxplot() +
  labs(title = "Illegal States Encountered by Disk Count", 
       x = "Disk Count", y = "Total Illegal States") +
  theme_minimal() +
  theme(legend.position = "none")
```

---

# Statistical Analysis

## Logistic Regression: Predicting Solvability

```{r logistic_regression}
# Fit logistic regression model
logit_model <- glm(is_solvable ~ disk_count + target_alteration_percent + 
                     illegal_formation_handler + ground_handler + duplicate_handler,
                   data = data, family = binomial)

# Summary
summary(logit_model)

# Odds ratios
exp(coef(logit_model)) %>%
  as.data.frame() %>%
  rownames_to_column(var = "Variable") %>%
  rename(Odds_Ratio = ".") %>%
  kable(caption = "Odds Ratios for Solvability", digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## ANOVA: Total Moves to Solve (Solvable Runs Only)

```{r anova_moves}
# Fit full factorial ANOVA model
anova_model <- aov(total_moves_to_solve ~ disk_count * target_alteration_percent * 
                     illegal_formation_handler * ground_handler * duplicate_handler,
                   data = data_solvable)

# ANOVA table
summary(anova_model)

# Type II Anova for unbalanced designs
Anova(anova_model, type = "II") %>%
  as.data.frame() %>%
  rownames_to_column(var = "Source") %>%
  kable(caption = "Type II ANOVA Table for Total Moves", digits = 4) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## ANOVA Diagnostics

```{r anova_diagnostics, fig.height=8}
par(mfrow = c(2, 2))
plot(anova_model)
par(mfrow = c(1, 1))
```

## Post-Hoc Comparisons: Tukey HSD

```{r tukey_hsd}
# Tukey HSD for main effects
tukey_disk <- TukeyHSD(aov(total_moves_to_solve ~ disk_count, data = data_solvable))
tukey_alteration <- TukeyHSD(aov(total_moves_to_solve ~ target_alteration_percent, data = data_solvable))
tukey_formation <- TukeyHSD(aov(total_moves_to_solve ~ illegal_formation_handler, data = data_solvable))

# Display results
tukey_disk$disk_count %>%
  as.data.frame() %>%
  rownames_to_column(var = "Comparison") %>%
  kable(caption = "Tukey HSD: Disk Count", digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

tukey_alteration$target_alteration_percent %>%
  as.data.frame() %>%
  rownames_to_column(var = "Comparison") %>%
  kable(caption = "Tukey HSD: Alteration Percentage", digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

tukey_formation$illegal_formation_handler %>%
  as.data.frame() %>%
  rownames_to_column(var = "Comparison") %>%
  kable(caption = "Tukey HSD: Formation Handler", digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Estimated Marginal Means

```{r emmeans}
# Calculate estimated marginal means
emm_disk <- emmeans(anova_model, ~ disk_count)
emm_alteration <- emmeans(anova_model, ~ target_alteration_percent)
emm_formation <- emmeans(anova_model, ~ illegal_formation_handler)

# Display results
summary(emm_disk) %>%
  as.data.frame() %>%
  kable(caption = "Estimated Marginal Means: Disk Count", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

summary(emm_alteration) %>%
  as.data.frame() %>%
  kable(caption = "Estimated Marginal Means: Alteration %", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

summary(emm_formation) %>%
  as.data.frame() %>%
  kable(caption = "Estimated Marginal Means: Formation Handler", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

# Two-Factor Interactions

## Disk Count × Alteration Rate

```{r interaction_disk_alt_analysis}
emm_interaction <- emmeans(anova_model, ~ disk_count * target_alteration_percent)

# Interaction plot
emmip(anova_model, disk_count ~ target_alteration_percent, CIs = TRUE) +
  labs(title = "Interaction Plot: Disk Count × Alteration Rate",
       x = "Target Alteration %", y = "Estimated Marginal Mean (Moves)",
       color = "Disk Count") +
  theme_minimal()
```

## Formation Handler × Alteration Rate

```{r interaction_formation_alt}
emmip(anova_model, illegal_formation_handler ~ target_alteration_percent, CIs = TRUE) +
  labs(title = "Interaction Plot: Formation Handler × Alteration Rate",
       x = "Target Alteration %", y = "Estimated Marginal Mean (Moves)",
       color = "Formation Handler") +
  theme_minimal()
```

---

# Alteration Analysis

## Actual vs Target Alteration Rate

```{r alteration_accuracy}
alteration_comparison <- data_solvable %>%
  group_by(target_alteration_percent) %>%
  summarise(
    mean_actual = mean(actual_alteration_percent, na.rm = TRUE),
    sd_actual = sd(actual_alteration_percent, na.rm = TRUE),
    n = n()
  ) %>%
  mutate(target = as.numeric(as.character(target_alteration_percent)))

alteration_comparison %>%
  kable(caption = "Actual vs Target Alteration Rates", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot
ggplot(alteration_comparison, aes(x = target, y = mean_actual)) +
  geom_point(size = 3, color = "steelblue") +
  geom_errorbar(aes(ymin = mean_actual - sd_actual, ymax = mean_actual + sd_actual), 
                width = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Target Alteration Rate",
       x = "Target Alteration %", y = "Actual Alteration %") +
  theme_minimal()
```

---

# Efficiency Metrics

## Overhead Analysis

```{r overhead_analysis}
# Calculate theoretical minimum moves for each disk count
theoretical_min <- data.frame(
  disk_count = c("3", "5", "7", "9"),
  theoretical_min_moves = c(7, 31, 127, 511)
)

overhead_data <- data_solvable %>%
  left_join(theoretical_min, by = "disk_count") %>%
  mutate(
    overhead = total_moves_to_solve - theoretical_min_moves,
    overhead_percent = (overhead / theoretical_min_moves) * 100
  )

overhead_summary <- overhead_data %>%
  group_by(disk_count, target_alteration_percent) %>%
  summarise(
    mean_overhead = mean(overhead, na.rm = TRUE),
    mean_overhead_pct = mean(overhead_percent, na.rm = TRUE),
    .groups = "drop"
  )

overhead_summary %>%
  kable(caption = "Overhead Analysis: Additional Moves Beyond Theoretical Minimum", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Visualization
ggplot(overhead_summary, aes(x = target_alteration_percent, y = mean_overhead_pct, 
                              color = disk_count, group = disk_count)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(title = "Overhead % by Disk Count and Alteration Rate",
       x = "Target Alteration %", y = "Mean Overhead %",
       color = "Disk Count") +
  theme_minimal()
```

---

# Conclusions

## Key Findings

1. **Disk Count Effect**: [To be filled based on results]
2. **Alteration Rate Impact**: [To be filled based on results]
3. **Handler Strategies**: [To be filled based on results]
4. **Interactions**: [To be filled based on results]

## Recommendations

1. [Based on statistical analysis]
2. [Based on efficiency metrics]
3. [Based on success rates]

## Future Work

- Investigate additional handler strategies
- Test with larger disk counts
- Explore non-uniform alteration distributions
- Develop predictive models for solvability

---

# Session Information

```{r session_info}
sessionInfo()
```
