---
title: "TwoFactor1"
author: "Luke Pepin"
date: "2025-10-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(agricolae)
library(emmeans)
library(DescTools)
library(pwr)
library(knitr)
library(kableExtra)
```

```{r data}
temp_levels <- c("Low", "Medium", "High")

data <- data.frame(
  material = rep(1:3, each = 12),
  temperature = rep(rep(temp_levels, each = 2), times = 3),
  value = c(
    #Material 1
    130, 155, 34, 40, 20, 70,
    74, 180, 80, 75, 82, 58,
    #Material 2
    150, 188, 136, 122, 25, 70,
    159, 126, 106, 115, 58, 45,
    #Material 3
    138, 110, 174, 120, 96, 104,
    168, 160, 150, 139, 82, 60
  )
)

data$material <- factor(data$material, labels = c("M1", "M2", "M3"))
data$temperature <- factor(data$temperature, levels = temp_levels)

str(data)

head(data)

ggplot(data, aes(x = material, y = value)) + 
  stat_summary(fun = mean, geom = "point", size = 2) +
  stat_summary(fun = mean, geom = "line", aes(group = 1), linewidth = 0.8) +
  labs(title = "Main Effect of Material", y = "Mean Value", x = "Material") + 
  theme_minimal(base_size = 10)

ggplot(data, aes(x = temperature, y = value)) + 
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun = mean, geom = "line", aes(group = 1), linewidth = 0.8) +
  labs(title = "Main Effect of Temperature", y = "Mean Value", x = "Material") + 
  theme_minimal(base_size = 10)
  
group_means <- data %>%
  group_by(material, temperature) %>%
  summarise(
    mean_value = mean(value),
    sd = sd(value),
    n = n(),
    se = sd / sqrt(n),
    .groups = 'drop'
  )

print(group_means)


ggplot(group_means, aes(x = temperature, y = mean_value, color = material, group = material)) + 
  geom_line(size = 1) + 
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.15) +
  labs(title = "Interaction Plot: Battery Life by Material and Temp.", y = "Mean Battery Life (hours)", x = "Temperature", color = " Material") + 
  theme_minimal(base_size = 10)
```

```{r analysis}
# Create visualization of main effects
# 1. Temperature effect
p1 <- ggplot(data, aes(x = temperature, y = value)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun = mean, geom = "line", aes(group = 1), size = 0.8) +
  labs(title = "Main Effect of Temperature", 
       y = "Battery Life (hours)", 
       x = "Temperature") +
  theme_minimal(base_size = 10)
print(p1)

# 2. Material effect
p2 <- ggplot(data, aes(x = material, y = value)) +
  stat_summary(fun = mean, geom = "point", size = 2) +
  stat_summary(fun = mean, geom = "line", aes(group = 1), size = 0.8) +
  labs(title = "Main Effect of Material", 
       y = "Battery Life (hours)", 
       x = "Material") +
  theme_minimal(base_size = 10)
print(p2)

# 3. Interaction between material and temperature
p3 <- ggplot(group_means, aes(x = temperature, y = mean_value, color = material, group = material)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.15) +
  labs(
    title = "Interaction Plot: Battery Life by Material and Temp.",
    x = "Temperature",
    y = "Mean Battery Life (hours)",
    color = "Material"
  ) +
  theme_minimal(base_size = 10)

# Fit two-way ANOVA model with interaction
# Model: Battery Life ~ Material * Temperature
model <- aov(value ~ material * temperature, data = data)

# Display ANOVA table with significance tests
summary(model)

# Store results and create effects plot
anova_sum <- summary(model)[[1]]
plot.design(value ~ material + temperature, data = data)

# Sample Size
k <- 9  # Number of groups (3 materials × 3 temperatures)
f <- 1  # Effect size (Cohen's f)
pwr_result <- pwr.anova.test(k = k, f = f, sig.level = 0.05, power = 0.9)
pwr_result

# Calculate statistical power from the experiment
# Define parameters
k <- 9 # total number of groups (3 plate materials * 3 temperatures)
f <- 1 # medium effect size; change as needed
# Calculate power
sample_calc <- pwr.anova.test(k = k, n = 4, f = f, sig.level = 0.05)$power
# Print results
sample_calc

# Tukey's HSD Test
tukey_result <- TukeyHSD(model)
print(tukey_result)
plot(tukey_result)

# Fisher's LSD Test
model <- aov(value ~ material * temperature, data = data)
lsd_resultM <- LSD.test(model, "material", p.adj = "none")
lsd_resultT <- LSD.test(model, "temperature", p.adj = "none")
print(lsd_resultM)
print(lsd_resultT)

# Assumption Tests
## Residuals vs Fitted
plot(model, which=1)
## Q-Q Residuals
plot(model, which=2)
```

# Conclusion

### **Validation**

The experiment was statistically robust. [cite_start]The sample size ($n=4$) was sufficient, and a post-hoc power calculation yielded a very high power of $0.98$, confirming the results are reliable[cite: 186, 209, 212].

---

### **Analysis and Recommendation**

[cite_start]The ANOVA found that **Material** ($p = 0.00198$), **Temperature** ($p < 0.0001$), and their **Interaction** ($p = 0.01861$) were all statistically significant[cite: 165].

[cite_start]The significant interaction means the effect of temperature depends on the material [cite: 84-101]. Key findings from the Tukey's HSD test are:
* [cite_start]**High temperature** performed significantly worse than "Low" or "Medium" for all materials[cite: 238, 239].
* [cite_start]The **M2/Low** (156.0 hours) and **M3/Medium** (146.0 hours) combinations produced the highest mean battery life[cite: 82].

**Recommendation:** To maximize battery life, use **Material 2 at a Low temperature** or **Material 3 at a Medium temperature**. High temperature settings should be avoided with any material.
